---
title: "MLandrum_Live_Session_4"
author: "Michael Landrum"
date: "9/23/2017"
output: html_document
---

# 1

You will need to install the fivethirtyeight package to run this script. After doing so, use that library to find all the data sets. I manually counted until the 18th data set which was named "college_recent_grads". Using vignette, I was able to track down the URL that had the article this information was used in.

The dimensions of the df are 173 x 21, and the column names are displayed below.

```{r FiveThirtyEight Data, echo=TRUE}
#install.packages("fivethirtyeight")
library(fivethirtyeight)
df <- college_recent_grads

# vignette("fivethirtyeight", package = "fivethirtyeight")
# http://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/

dim(df)
colnames(df)
```

# 2

We are told to print the column names again, so I did, and we can see that it is confirmed there are 21 columns.

I'm sure there was a quicker way to finding the frequency of each category but I wanted to practice making loops. I created two empty vectors, one a character vector and one numeric. I iterated through each string in df$major_category and checked to see if it was already in the major vector. If it was, I added to the count of that major. If it was not, I created a new element in the major vector, and set the count to 1.

I used sum(count) to make sure it had 173 like our original data frame.

``` {r Data Summary a-b, echo=TRUE}
#colnames(df)
ncol(df)

major <- character()
count <- numeric()
for(i in df$major_category){
  if( i %in% major ){
    x <- match(i, major)
    count[x] <- count[x] + 1
  } else{
    y <- length(major) + 1
    major[y] <- i
    count[y] <- 1
  }
}

major_count <- data.frame( major, count )
major_count
#sum(count)

```

A bar plot of the data is shown below. I removed the x-label due to it covering some of the majors. But you can unhide the comment to include it.

If you'd like to have a csv of this data, uncomment the final line.

``` {r Data Summary c-d, echo=TRUE}
par(las=2)
barplot(height = major_count$count, names.arg = major_count$major,
        main = "The Economic Guide to Picking a College Major",
        col = 2, ylab = "Frequency", 
        #xlab = "College Majors"
        )

#write.csv(major_count, file = "major_count.csv", row.names=FALSE)
```

3. My github repo can be found at https://github.com/mlandrum22/smuMSDS6306.git 