---
title: "SP500"
author: "Michael Landrum"
date: "11/8/2017"
output: html_document
---

## Unit 11: Modeling Financial Data

Uncomment and install packages if you don't have it

``` {r install, echo=TRUE, message=FALSE}
#install.packages("tseries")
library(tseries)
```

### Description: Please fill out that this is for educational purposes and talk about what it's for!

Download the data of SP500 '^gspc'.

``` {r download, echo=TRUE, message=FALSE, warning=FALSE}
SNPdata <- get.hist.quote("^gspc",quote="Close")
```

Calculate the log returns, which is the subtraction of log(lag(SNPdata)) and log(SNPdata)

``` {r logReturns, echo=TRUE}
SNPret <- log(lag(SNPdata)) - log(SNPdata)
```

Calculate volatility measure that is to multiply sd(SNPret),sqrt(250), 100

``` {r volatlity, echo=TRUE}
SNPvol <- sd(SNPret) * sqrt(250) * 100
```

Define getVol function for volatility
``` {r getVol, echo=TRUE}
getVol <- function(d, logrets) {
	var = 0
	lam = 0
	varlist <- c()

	for (r in logrets) {
		lam = lam*(1 - 1/d) + 1
	  var = (1 - 1/lam)*var + (1/lam)*r^2
		varlist <- c(varlist, var)
	}

	sqrt(varlist)
}

```

Calculate volatility over entire length of series for various three different decay factors: 10, 30, 100

``` {r volests, echo=TRUE}
volest <- getVol(10, SNPret)
volest2 <- getVol(30, SNPret)
volest3 <- getVol(100, SNPret)
```

Plot the results, overlaying the volatility curves on the data, just as was done in the S&P example.


``` {r plots, echo=TRUE}
plot(volest,type="l")
lines(volest2,type="l",col="red")
lines(volest3,type="l",col="blue")
```